Plant Metagenomic Barcoding using Angiosperms353 of Corbiculae from wild Bumble Bees

### 2023-03-07 script/manuscript_MEE.Rmd
added citations to first paragraph

An enormous amount of Earths biodiversity may be attributed to the interactions between species (@soltis2019darwin, @futuyma2009macroevolution, @voje2015role, @weber2017evolution, @hembry2020ecological). These interactions not only leading to the the origin of many species but are essential to the maintenance of virtually all ecosystems (CITE). In order to understand and rationally conserve both the species and the ecosystems which their interactions compose - Darwins 'Entangled Bank' integrative approaches, with the potential for scaling are required immediately (@darwin2004origin, @thompson1994coevolutionary, @banerjee2022environmental). Major limitations, imposed by a lack of taxonomic expertise, impede our initial ability to identify organisms let alone more complex phenomena (@hebert2003biological). The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a role they are generally pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic). The ability to identify fragments of organisms (e.g. leaf tissue) increases our ability to understand the interactions of a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g. those for restoration processes (@banerjee2022environmental, @johnson2023environmental).

@@

  Currently the largest plant systematic endeavor ever undertaken,by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).  This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000 represented species, using the popular Angiosperms353 (A353) probes, which includes 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal). These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding. The A353 probes are currently being used in many other plant phylogenetic studies increasing the sampling depth of many clades  (@baker2021exploring). Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute many more species by 2030 (@cheng2018tenkp). Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high numbers of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@darwin2022project, @lewin2022biogenome). These data will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling). However, the application of metabarcoding still faces challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.

>>>>

An enormous amount of Earths biodiversity may be attributed to the interactions between species (@soltis2019darwin, @futuyma2009macroevolution, @voje2015role, @weber2017evolution, @hembry2020ecological). These interactions not only lead to the the origin of many species but appear essential to the maintenance of virtually all ecosystems (@agrawal2007filling, @valiente2015beyond, @bascompte2006asymmetric). In order to understand and rationally conserve both the species and the ecosystems which their interactions compose - Darwins 'Entangled Bank'- integrative approaches, with the potential for scaling are required immediately (@darwin2004origin, @thompson1994coevolutionary, @agrawal2007filling, @banerjee2022environmental, @blanchet2020co, @jordano2016sampling). Major limitations, imposed by a lack of taxonomic expertise, impede our initial ability to identify organisms let alone more complex phenomena (@hebert2003biological). The lack of an ability to identify whole organisms, especially those from diverse clades where species are oftentimes delineated along ecological lines - or on occasion cryptic, lessens the utility of them to serve as bioindicators; a role they are typically pre-disposed to (@Gage2013HistoricalRO, @banerjee2022environmental, @janzen2017nuclear, @oliver2009cryptic). The ability to identify fragments of organisms (e.g. leaf tissue) increases our ability to understand the interactions of  not only entire ecosystems but also a focal, generally rare and hence difficult to detect, organism with their surrounding; allowing for the most precise allocation of conservation decisions and funds e.g. those for restoration processes (@banerjee2022environmental, @johnson2023environmental).

@@

  Currently the largest plant systematic endeavor ever undertaken,by the Royal Botanic Gardens Kew, the Plant and Fungal Tree of Life (PAFTOL) is approaching completion (@baker2021PAFTOL).  This data set will contain hybridization capture (Hyb-Seq) data from at least one species in each genus of the plant kingdom, 14,000 represented species, using the popular Angiosperms353 (A353) probes, which includes 353 single-copy orthologous loci, (@baker2021PAFTOL, @johnson2019universal). These publicly available data serve to provide a taxonomically comprehensive backbone for plant metabarcoding. Data from the 10kP project, which seeks to develop reference genomes from a phylogenetically diverse suite of 10,000 plant species, will contribute many more species by 2030 (@cheng2018tenkp). Similar projects such as the ‘Darwin Tree of Life’ which will sequence all described taxa in Britain and Ireland, seek to sequence high numbers of genomes in geographic regions will contribute data sets applicable to enormous spatial domains (@darwin2022project, @lewin2022biogenome). These data will promote the ability to apply metabarcoding to resolve a diverse array of questions relevant to theoretical and applied ecology (@kress2017plant, @hollingsworth2016telling). However, the application of metabarcoding still faces challenges relating to the enormity of the genomic data sets and the computational power required to process sequence data.
